Evolution of Methodologies in Delivering Reliable Software in the European Scientific Arena

### 2018-06-05 paper-ieee-trans.tex

% trial .tex file %

@@

The E--Infrastructure for Testing, Integration and Configuration of Software
\cite{cordis:etics, cordis:etics2} ({\sl ETICS}, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as {\sl EGEE} and its {\sl gLite} middleware. The framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto--generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The {\sl ETICS} framework was the first automated
service for delivering quality software products in distributed environments like
the Grids.

@@

For each software characteristic, a set of associated
metrics and Key Performance Indicators (KPIs) were identified and defined in
detail in the {\sl EMI} Metrics Specification \cite{emi-metrics}. The project
leveraged the {\sl ETICS} service for the development, continuous integration and release management,
as well as for metric tracking, making queries on the collected data to display
them through the chart generation framework.

@@

\item \textit{Documentation}.
The SQA criteria set the path for the adoption of the developed software by
defining the documentation content, according to the target audience. This requirement also
promotes the automation, both in terms of documentation creation and service deployment. The
\textit{documentation is treated as code}, using a markup language, automatically rendered and
uploaded to online repositories \cite{indigo-gitbook}. Thus documentation is portable and
human--reviewed, using the same workflow as the code does, validating the changes before being
updated into the production repository.

\item \textit{Automated Deployment}.
To lower the barriers of software adoption, the SQA criteria require the automated
deployment of the products delivered as part of the catalogue. Automation in this context is
tackled using a configuration management tool. These tools allow the adoption of
\textit{Infrastructure as Code} (IaC), managing the component's deployment through declarative
definitions. The definition files appear as additional source of documentation
-- self--documenting code -- as they sequentially guide the component's deployment process in multiple
platforms. {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such
as Ansible \cite{indigo-galaxy} and Puppet \cite{indigo-puppet}. A representative example of
such contribution are the \textit{50 roles} developed from scratch and currently hosted in
the Ansible Galaxy portal \cite{indigo-galaxy}. Fig.~\ref{fig:fig_confman}
shows the number of products that offer an automated means for deployment.
It shows an increase in the adoption of such tools leading to the {\sl INDIGO--1} release
(dark blue points and trend line), as well as to the {\sl INDIGO--2} release
(light cream points and trend line). The rate of adoption is lower during the weeks
before the second release because a significant fraction of the products had already adopted it
previously to the first release.
\end{itemize}

@@

\textit{Development metrics} are obtained programmatically from several sources, namely GitHub
API and Jenkins service, and graphically displayed as GitHub pages using GrimoireLab
framework \cite{grimoirelab}. Per--component weekly reports, including the SQA requirement
fulfilment, are issued and individually discussed through the different communication channels
with the development teams. Issue tracking metrics and KPIs are an essential piece of information
of both feature addition and defect solving, useful to detect and fix misbehaviours while in the
development phase.

\textit{Release metric} sources are the online repository servers, namely the
Linux package \cite{indigo-pkg-repo} and DockerHub \cite{indigo-dockerhub} repositories. Jenkins
server also contains valuable release data as it is the service where the packages
and containers are being built before being uploaded to the online repositories. Release
KPIs primarily focus on the frequency and efficiency -- mainly rollbacks -- statistics.

\textit{Maintenance and user support metrics} are key for continuously improving the response
to issues reported by external users. Feedback is collected from outside helpdesks, such as
EGI's GGUS \cite{ggus}, and the GitHub Issues tracker.

@@

The practical implementation of the CI scenario leveraged on tightly integrated open
source tools such as GitHub \cite{github}, as the online source code repository,

% Comment
%%not a sentence, again
Jenkins \cite{jenkins}, the event--response CI system, and Docker container
provisioning to provide instant computing power needed for executing the pipeline jobs.
The project defined a source code contribution workflow, based on GitHub Pull Requests
(PRs), where each change automatically triggers in Jenkins the associated quality and
metric tracking checks on PR creation or update. The set of required checks contained
in the CI pipeline, report their exit status back to GitHub preventing the change to be
approved if they are not successfully executed. As each change is validated, the chances
of early detection of defects increases. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\bibitem{h2020}
\emph{HORIZON 2020}, The EU Framework Programme for Research and Innovation [Online]. Available: https://ec.europa.eu/programmes/horizon2020/ .

\bibitem{agile-manifesto}
\emph{Manifesto for Agile Software Development} [Online]. Available: http://agilemanifesto.org .

\bibitem{zhu}
L. Zhu and L. Bass and G. Champlin--Scharff, \emph{DevOps and Its Practices},
IEEE Software, vol. 33, no. 3, pp. 32--34, May--June 2016.

\bibitem{cordis:datagrid}
\emph{DataGrid project}, European Community Research and Development
Information Service (CORDIS) [Online]. Available: http://cordis.europa.eu/project/rcn/53665\_en.html .

\bibitem{gagliardi}
Gagliardi F., Jones B., Reale M., Burke S., \emph{European DataGrid Project: Experiences of Deploying a Large Scale Testbed for E--science Applications},
in Calzarossa M.C., Tucci S. (eds) Performance Evaluation of Complex Systems: Techniques and Tools, Lecture Notes in Computer Science, vol 2459. Springer, 2002.

\bibitem{globus}
I. Foster and C. Kesselman, \emph{Globus: a Metacomputing Infrastructure
Toolkit}, International Journal of Supercomputer Applications, vol. 11, no. 2,
pp. 115--128, 1997.

\bibitem{datagrid}
L. Momtahan and A. Martin, \emph{e--Science Experiences: Software Engineering
Practice and the EU DataGrid}, in Proc. Asia--Pacific Software Engineering
Conference, Gold Coast, Queensland, Australia, pp. 269-275, IEEE Press,
4-6 Dec. 2002.

\bibitem{agile}
T. Dingsoyr, \emph{A decade of agile methodologies: Towards explaining agile
software development} in The Journal of Systems and Software, Elsevier Inc,
2012.

\bibitem{cmm}
Paulk et al., \emph{Capability maturity model for software}, Software
Engineering Institute, report CMU/SEI-93-TR-24, ESC--TR-93-177, Feb. 1993.

\bibitem{cordis:egee}
\emph{Enabling Grids for E--sciencE (EGEE)} project, European Community
Research and Development Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/80149\_en.html .

\bibitem{cordis:egee2}
\emph{Enabling Grids for E--sciencE--II (EGEE--II)} project, European Community
Research and Development Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/99189\_en.html .

\bibitem{cordis:egee3}
\emph{Enabling Grids for E--sciencE--III (EGEE--III)} project, European Community
Research and Development Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/87264\_en.html .

\bibitem{egee:acceptance-criteria}
\emph{Definition and Documentation of the Revised Software Life--Cycle Process},
Milestone MSA3.4.2, 2010, EGEE--III project [Online]. Available: https://edms.cern.ch/ui/file/1062487/2/EGEE--III--MSA3.4.2-1062487-v1\_4.pdf .

\bibitem{glite}
E. Laure et al, \emph{Programming the Grid with gLite}, in Jin, H., Reed, D.A.,
Jiang, W. (eds.) Computational Methods in Science and Technology, vol. 12(1),
pp. 33--45, Scientific Publishers OWN, 2006.

\bibitem{condor}
D. Thain, T. Tannenbaum, M. Livny, \emph{Condor and the Grid}, in Grid
Computing: Making the Global Infrastructure a Reality, Chapter 11, pp. 63--70,
Eds. John Wiley \& Sons Inc., 2002.

\bibitem{cordis:etics}
\emph{E--Infrastructure for Testing, Integration and Configuration of Software
(ETICS)} project, European Community Research and Development Information
Service (CORDIS) [Online]. Available: http://cordis.europa.eu/project/rcn/80138\_en.html .

\bibitem{cordis:etics2}
\emph{E--Infrastructure for Testing, Integration and Configuration of Software -
Phase 2 (ETICS 2)} project, European Community Research and Development
Information Service (CORDIS), [Online]. Available: http://cordis.europa.eu/project/rcn/86604\_en.html .

\bibitem{etics}
A. Di Meglio, M.-E.B??gin, P. Couvares, E. Ronchieri, E. Takacs, \emph{ETICS:
the international software engineering service for the Grid}, in Journal of
Physics: Conference Series, Vol. 119, N. 4, IOP Publishing Ltd, 2008.

\bibitem{cordis:emi}
\emph{European Middleware Initiative (EMI)} project, European Community
Research and Development Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/95311\_en.html .

\bibitem{iso-9126}
\emph{ISO/IEC 9126 Software Engineering - Product Quality}, International
Organization for Standardization [Online]. Available: https://www.iso.org/standard/22749.html .

\bibitem{emi-metrics}
\emph{EMI Metrics Specification} [Online]. Available: https://goo.gl/CCtY7x .

\bibitem{emi-quality-model}
\emph{EMI Quality Model} [Online]. Available: https://goo.gl/LdS6fL .

\bibitem{cordis:egi-inspire}
\emph{European Grid Initiative: Integrated Sustainable Pan--European
Infrastructure for Researchers in Europe (EGI--InSPIRE)} project, European
Community Research and Development Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/95923\_en.html .

\bibitem{egi-ops}
\emph{EGI Operations Portal} [Online]. Available:
https://operations-portal.egi.eu/ .

\bibitem{ggus}
\emph{GGUS} EGI Helpdesk [Online]. Available:
https://ggus.eu/ .

\bibitem{gocdb}
\emph{GOCDB} EGI Grid Configuration Repository [Online]. Available:
https://goc.egi.eu/ .

\bibitem{egi-qc}
\emph{EGI Quality Criteria} [Online]. Available:
https://egi-qc.github.io/ .

\bibitem{github}
\emph{GITHUB} [Online]. Available:
https://github.com/ .

\bibitem{jenkins}
\emph{Jenkins} Continuous Integration System [Online]. Available:
https://jenkins-ci.org/ .

\bibitem{mario}
M. David et al, \emph{Validation of Grid Middleware for the European Grid
Infrastructure}, in Journal of Grid Computing, vol. 12, issue 3, pp. 543--558,
Springer, 2014.

\bibitem{cordis:egi-engage}
\emph{Engaging the EGI Community towards an Open Science Commons (EGI--ENGAGE)}
project, European Community Research and Development Information Service
(CORDIS), [Online]. Available:  http://cordis.europa.eu/project/rcn/194937\_en.html .

\bibitem{cordis:indigo}
\emph{INtegrating Distributed data Infrastructures for Global ExplOitation
(INDIGO--DataCloud)} project, European Community Research and Development
Information Service (CORDIS) [Online]. Available:
http://cordis.europa.eu/project/rcn/194882\_en.html .

\bibitem{indigo-dockerhub}
\emph{INDIGO--DataCloud DockerHub repository} [Online]. Available:
https://hub.docker.com/u/indigodatacloud .

\bibitem{indigo-d31}
\emph{Initial Plan for WP3} INDIGO--DataCloud Deliverable 3.1 [Online]. Available:
https://www.indigo-datacloud.eu/documents/initial-plan-wp3-d31 .

\bibitem{indigo-jenkins}
\emph{INDIGO--DataCloud Jenkins CI service} [Online]. Available:
https://jenkins.indigo-datacloud.eu:8080/ .

\bibitem{indigo-github}
\emph{INDIGO--DataCloud GitHub Source Code repository} [Online]. Available:
https://www.github.com/indigo-dc .

\bibitem{indigo-gitbook}
\emph{INDIGO--DataCloud GitBook Documentation repository} [Online]. Available:
https://www.gitbook.com/@indigo-dc .

\bibitem{indigo-puppet}
\emph{INDIGO--DataCloud PuppetForge repository} [Online]. Available:
https://forge.puppet.com/indigodc .

\bibitem{indigo-galaxy}
\emph{INDIGO--DataCloud Ansible Galaxy repository} [Online]. Available:
https://galaxy.ansible.com/indigo-dc/ .

\bibitem{grimoirelab}
\emph{GrimoireLab} [Online]. Available: http://grimoirelab.github.io/ .

\bibitem{indigo-pkg-repo}
\emph{INDIGO--DataCloud Linux Package repository} [Online]. Available: https://jenkins.indigo-datacloud.eu/ .

\bibitem{disvis} G.C.P. Van Zundert, A.M.J.J. Bonvin,
DisVis: quantifying and visualizing the accessible interaction space of distance
restrained biomolecular complexes, Bioinformatics 31 (2015) 3222--3224 .

\bibitem{powerfit} G.C.P. Van Zundert, A.M.J.J. Bonvin,
Fast and sensitive rigid--body fitting into cryo--EM density maps with PowerFit,
AIMS Biophys. 2 (2015) 73--87.

>>>>

@@

The E--Infrastructure for Testing, Integration and Configuration of Software~\cite{etics}
({\sl ETICS}, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as {\sl EGEE} and its {\sl gLite} middleware. The framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto--generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The {\sl ETICS} framework was the first automated
service for delivering quality software products in distributed environments like
the Grids.

@@

For each software characteristic, a set of associated
metrics and Key Performance Indicators (KPIs) were identified and defined in
detail in the {\sl EMI} Metrics Specification \cite{emi-quality-model}. The project
leveraged the {\sl ETICS} service for the development, continuous integration and release management,
as well as for metric tracking, making queries on the collected data to display
them through the chart generation framework.

@@

\item \textit{Documentation}.
The SQA criteria set the path for the adoption of the developed software by
defining the documentation content, according to the target audience. This requirement also
promotes the automation, both in terms of documentation creation and service deployment. The
\textit{documentation is treated as code}, using a markup language, automatically rendered and
uploaded to online repositories\footnote{\url{https://www.gitbook.com/@indigo-dc}}.
Thus documentation is portable and
human--reviewed, using the same workflow as the code does, validating the changes before being
updated into the production repository.

\item \textit{Automated Deployment}.
To lower the barriers of software adoption, the SQA criteria require the automated
deployment of the products delivered as part of the catalogue. Automation in this context is
tackled using a configuration management tool. These tools allow the adoption of
\textit{Infrastructure as Code} (IaC), managing the component's deployment through declarative
definitions. The definition files appear as additional source of documentation
-- self--documenting code -- as they sequentially guide the component's deployment process in multiple
platforms. {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such
as Ansible
and Puppet\footnote{\url{https://forge.puppet.com/indigodc}}. A representative example of
such contribution are the \textit{50 roles} developed from scratch and currently hosted in
the Ansible Galaxy portal\footnote{\url{https://galaxy.ansible.com/indigo-dc/}}. Fig.~\ref{fig:fig_confman}
shows the number of products that offer an automated means for deployment.
It shows an increase in the adoption of such tools leading to the {\sl INDIGO--1} release
(dark blue points and trend line), as well as to the {\sl INDIGO--2} release
(light cream points and trend line). The rate of adoption is lower during the weeks
before the second release because a significant fraction of the products had already adopted it
previously to the first release.
\end{itemize}

@@

\textit{Development metrics} are obtained programmatically from several sources, namely GitHub
API\footnote{\url{https://developer.github.com/v3/}.}
and Jenkins\footnote{\url{https://jenkins.io/}} service, and graphically displayed as GitHub
pages using GrimoireLab framework\footnote{\url{http://grimoirelab.github.io/}}.
Per--component weekly reports, including the SQA requirement
fulfilment, are issued and individually discussed through the different communication channels
with the development teams. Issue tracking metrics and KPIs are an essential piece of information
of both feature addition and defect solving, useful to detect and fix misbehaviours while in the
development phase.

\textit{Release metric} sources are the online repository servers, namely the
Linux package\footnote{\url{https://jenkins.indigo-datacloud.eu/}}
and DockerHub\footnote{\url{https://hub.docker.com/u/indigodatacloud}} repositories. Jenkins
server also contains valuable release data as it is the service where the packages
and containers are being built before being uploaded to the online repositories. Release
KPIs primarily focus on the frequency and efficiency -- mainly rollbacks -- statistics.

\textit{Maintenance and user support metrics} are key for continuously improving the response
to issues reported by external users. Feedback is collected from outside helpdesks, such as
EGI's GGUS~\cite{ggus}, and the GitHub Issues tracker.

@@

The practical implementation of the CI scenario leveraged on tightly integrated open
source tools such as GitHub\footnote{\url{https://github.com/}}, as the online source code repository,

% Comment
%%not a sentence, again
Jenkins, the event--response CI system, and Docker container
provisioning to provide instant computing power needed for executing the pipeline jobs.
The project defined a source code contribution workflow, based on GitHub Pull Requests
(PRs), where each change automatically triggers in Jenkins the associated quality and
metric tracking checks on PR creation or update. The set of required checks contained
in the CI pipeline, report their exit status back to GitHub preventing the change to be
approved if they are not successfully executed. As each change is validated, the chances
of early detection of defects increases. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\bibitem{h2020}
\emph{Horizon 2020} -- European Commission.
[Online] Available: \url{https://ec.europa.eu/programmes/horizon2020/}
[Accessed 4 Jun. 2018].

\bibitem{agile-manifesto}
Beck, K., Beedle, et al., (2001).
Manifesto for Agile Software Development.
[Online] Available: \url{http://www.agilemanifesto.org/}
[Accessed 4 Jun. 2018].

\bibitem{zhu}
L. Zhu and L. Bass and G. Champlin--Scharff,
DevOps and Its Practices,
\emph{IEEE Software},
vol. 33, n. 3, pp. 32--34, 2016.

\bibitem{cordis:datagrid}
P. Kunszt,
European DataGrid project: status and plans,
\emph{Nucl. Instr. Meth. Phys. Res. A},
vol. 502, n. 2, pp. 376--381, 2003.
%\emph{DataGrid project}, European Community Research and Development Information Service (CORDIS).
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/53665\_en.html}
%[Accessed 4 Jun. 2018].

\bibitem{gagliardi}
Gagliardi F., Jones B., Reale M., Burke S.,
European DataGrid Project: Experiences of Deploying a Large Scale Testbed for E--science Applications,
in \emph{Performance Evaluation of Complex Systems: Techniques and Tools, LNCS},
vol. 2459, 2002.

\bibitem{globus}
I. Foster and C. Kesselman,
Globus: a Metacomputing Infrastructure Toolkit,
\emph{Int. J. Supercomput. Appl.},
vol. 11, no. 2, pp. 115--128, 1997.

\bibitem{datagrid}
L. Momtahan and A. Martin,
e--Science Experiences: Software Engineering Practice and the EU DataGrid,
in \emph{Ninth Asia--Pacific Software Engineering Conference},
pp. 269-275, 2002.

\bibitem{agile}
Torgeir Dingsoyr, Sridhar Nerur, VenuGopal Balijepally, Nils Brede Moe,
A decade of agile methodologies: Towards explaining agile software development,
\emph{J. Syst. Softw.},
vol. 85, Issue 6, pp. 1213--1221, June 2012.

\bibitem{cmm}
Paulk M., Curtis B., Chrissis M., and Weber V. C.,
Capability maturity model for software,
\emph{Software Engineering Institute},
report CMU/SEI-93-TR-24, ESC--TR-93-177, 1993.

\bibitem{cordis:egee}
F. Gagliardi and M. E. B\'{e}gin,
EGEE -- providing a production quality grid for e-science,
in \emph{2005 IEEE Inter. Symp. on Mass Storage Systems and Technology},
pp. 88-92, 2005.
%\emph{Enabling Grids for E--sciencE (EGEE)} project, European Community Research and
%Development Information Service (CORDIS). 
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/80149\_en.html}
%[Accessed 4 Jun. 2018].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TOBE checked if there are papers on these
\bibitem{cordis:egee2}
\emph{Enabling Grids for E--sciencE--II (EGEE--II)} project, European Community Research and
Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/99189\_en.html}
[Accessed 4 Jun. 2018].

\bibitem{cordis:egee3}
\emph{Enabling Grids for E--sciencE--III (EGEE--III)} project, European Community
Research and Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/87264\_en.html}
[Accessed 4 Jun. 2018].
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibitem{egee:acceptance-criteria}
\emph{Definition and Documentation of the Revised Software Life--Cycle Process},
Milestone MSA3.4.2, 2010, EGEE--III project.
[Online] Available: \url{https://edms.cern.ch/ui/file/1062487/2/EGEE-III-MSA3.4.2-1062487-v1_4.pdf}
[Accessed 4 Jun. 2018].

\bibitem{glite}
E. Laure et al.,
Programming the Grid with gLite,
\emph{Computational Methods in Science and Technology},
vol. 12(1), pp. 33--45, 2006.

\bibitem{condor}
D. Thain, T. Tannenbaum, M. Livny,
Condor and the Grid,
in \emph{Grid Computing: Making the Global Infrastructure a Reality},
Chapter 11, pp. 63--70, 2002.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TOBE checked if there are papers on these
%\bibitem{cordis:etics}
%\emph{E--Infrastructure for Testing, Integration and Configuration of Software (ETICS)}
%project, European Community Research and Development Information Service (CORDIS).
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/80138\_en.html}
%[Accessed 4 Jun. 2018].

%\bibitem{cordis:etics2}
%\emph{E--Infrastructure for Testing, Integration and Configuration of Software -
%Phase 2 (ETICS 2)} project, European Community Research and Development
%Information Service (CORDIS).
%[Online]. Available: \url{http://cordis.europa.eu/project/rcn/86604\_en.html}
%[Accessed 4 Jun. 2018].
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibitem{etics}
A. Di Meglio, et al.,
ETICS: the international software engineering service for the grid,
in \emph{J. Phys.: Conf. Ser.},
vol. 119, n. 4, 042010, 2008.

\bibitem{cordis:emi}
C. Aiftimiei, et al.,
Towards next generations of software for distributed infrastructures: The European Middleware Initiative,
in \emph{2012 IEEE 8th Inter. Conf. on E-Science},
pp. 1--10, 2012.
%\emph{European Middleware Initiative (EMI)} project, European Community
%Research and Development Information Service (CORDIS).
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/95311\_en.html}
%[Accessed 4 Jun. 2018].

\bibitem{iso-9126}
\emph{ISO/IEC 9126 Software Engineering - Product Quality},
International Organization for Standardization.
[Online] Available: \url{https://www.iso.org/standard/22749.html}
[Accessed 4 Jun. 2018].

\bibitem{emi-quality-model}
M Alandes, et al.,
Experiences with Software Quality Metrics in the EMI middleware,
in \emph{J. Phys.: Conf. Ser.}
vol. 396, n. 5, 052003, 2012.

\bibitem{cordis:egi-inspire}
I. Campos Plasencia,
EGI.eu the European grid initiative.
in \emph{4 th Iberian Grid Infrastructure Conf. Proc.},
pp. 5--15, 2010.
%\emph{European Grid Initiative: Integrated Sustainable Pan--European
%Infrastructure for Researchers in Europe (EGI--InSPIRE)} project, European
%Community Research and Development Information Service (CORDIS).
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/95923\_en.html}
%[Accessed 4 Jun. 2018].

\bibitem{egi-ops}
H. Cordier, et al.,
From EGEE Operations Portal towards EGI Operations Portal,
in \emph{Data Driven e-Science (ISGC2010)},
pp. 129--140, 2011.
%\emph{EGI Operations Portal}.
%[Online] Available: \url{https://operations-portal.egi.eu/}
%[Accessed 4 Jun. 2018].

\bibitem{ggus}
T. Antoni, et al.,
Global grid user support—building a worldwide distributed user support infrastructure,
\emph{J. Phys.: Conf. Ser.},
vol. 119, n. 5, 052002, 2008.
%\emph{GGUS} EGI Helpdesk.
%[Online] Available: \url{https://ggus.eu/}
%[Accessed 4 Jun. 2018].

\bibitem{gocdb}
G. Mathieu and J. Casson,
GOCDB4, a New Architecture for the European Grid Infrastructure,
in \emph{Data Driven e-Science (ISGC2010)},
pp. 163--174, 2011.
%\emph{GOCDB} EGI Grid Configuration Repository.
%[Online] Available: \url{https://goc.egi.eu/}
%[Accessed 4 Jun. 2018].

\bibitem{egi-qc}
\emph{EGI Quality Criteria}.
[Online] Available: \url{https://egi-qc.github.io/}
[Accessed 4 Jun. 2018].

\bibitem{mario}
M. David et al.,
Validation of Grid Middleware for the European Grid Infrastructure,
\emph{Journal of Grid Computing},
vol. 12, issue 3, pp. 543--558, 2014.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TOBE checked if there are papers on these
\bibitem{cordis:egi-engage}
\emph{Engaging the EGI Community towards an Open Science Commons (EGI--ENGAGE)}
project, European Community Research and Development Information Service (CORDIS).
[Online] Available: \url{http://cordis.europa.eu/project/rcn/194937\_en.html}
[Accessed 4 Jun. 2018].
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibitem{cordis:indigo}
Davide Salomoni, et al.,
INDIGO-DataCloud: {A} data and computing platform to facilitate seamless access to e-infrastructures,
\emph{CoRR},
vol. abs/1711.01981, 2018,
[Online] Available: \url{http://arxiv.org/abs/1711.01981}.
%\emph{INtegrating Distributed data Infrastructures for Global ExplOitation
%(INDIGO--DataCloud)} project, European Community Research and Development
%Information Service (CORDIS).
%[Online] Available: \url{http://cordis.europa.eu/project/rcn/194882\_en.html}
%[Accessed 4 Jun. 2018].

\bibitem{indigo-d31}
J. Gomes, et al.,
\emph{Initial Plan for WP3} INDIGO--DataCloud Deliverable 3.1.
[Online] Available: \url{https://www.indigo-datacloud.eu/documents/initial-plan-wp3-d31}
[Accessed 4 Jun. 2018].

\bibitem{disvis}
G.C.P. Van Zundert, A.M.J.J. Bonvin,
DisVis: quantifying and visualizing the accessible interaction space of distance
restrained biomolecular complexes,
\emph{Bioinformatics}, vol. 31, n. 19, pp. 3222--3224, 2015.

\bibitem{powerfit}
G.C.P. Van Zundert, A.M.J.J. Bonvin,
Fast and sensitive rigid--body fitting into cryo--EM density maps with PowerFit,
\emph{AIMS Biophys.}, vol. 2, n. 20150273, 73--87, 2015.
